---
title: "STAT4355HW5"
subtitle: This is an R script with the purpose of running multiple linear regression
  on football data
output:
  html_notebook: default
  pdf_document: default
---


### (a)
```{r}
#load the data
ftbl <- read.csv(file = 'football.csv')
#9 scatterplots for number of games won (y) against other vars
plot(ftbl$x1, ftbl$y)
plot(ftbl$x2, ftbl$y)
plot(ftbl$x3, ftbl$y)
plot(ftbl$x4, ftbl$y)
plot(ftbl$x5, ftbl$y)
plot(ftbl$x6, ftbl$y)
plot(ftbl$x7, ftbl$y)
plot(ftbl$x8, ftbl$y)
plot(ftbl$x9, ftbl$y)
```
### (b)
```{r}
#build model with most associated vars
lm1 <- lm(y~x1+x2+x5+x7+x8, data = ftbl)
#model summary
summary(lm1)
sigma(lm1)^2
```
Fitted linear model:

y-hat = -0.0516532 + 0.0008480x1 + 0.0034689x2 + 0.0103765x5 + 0.1336928x7 + -0.0047048x8 

(i)
(σ^)^2 = 3.126997
(ii)
R^2 = 0.7896
(iii)
adjusted R^2 = 0.7418

### (c)
βˆ1: For a one unit increase in rushing yardage, number of games won is associated with a 0.0008480 increase on average.

βˆ2: For a one unit increase in passing yardage, number of games won is associated with a 0.0034689 increase on average.

βˆ5: For a one unit increase in turnover differential, number of games won is associated with a 0.0103765 increase on average.

βˆ7: For a one unit increase in percentage of rushing plays, number of games won is associated with a 0.1336928 increase on average.

βˆ8: For a one unit increase in  opponents’ rushing yardage, number of games won is associated with a 0.0047048 decrease on average.

(σ^)^2: The model on average has a squared error of 3.126997.

### (d)
```{r}
#anova from pg.46
fit <- lm(y~x1+x2, data=ftbl)
anova(fit)

#Source of Variation   SS      DF      MS      F0 
#Regression          198.41    2     99.205   19.2926
#Residual            128.553   25    5.14212 
#Total               326.963   27   
```
ANOVA Table commented in code above^

### (e)
```{r}
#anova from pg.47
fit0<-lm(y~1, data=ftbl)
anova(fit0, fit)

#Source of Variation   SS      DF      MS      F0 
#Regression          198.41    2     99.205   19.293
#Residual            128.55    25     5.142 
#Total               326.96    27    
```
ANOVA Table commented in code above^

### (f)
```{r}
#regression significance test at level of 0.05
qf(0.05, 2, 25, lower.tail=FALSE)
```
The two ANOVA tables are the same.

### (g)
```{r}
#Test the value of passing yardage given all the other 4 predictors at significance level of 0.05
reduced <- lm(y~x1+x5+x7+x8, data = ftbl)
anova(reduced, lm1)
```
i. 

H0: βˆ2 = 0

H1: βˆ2 != 0

ii. F-distribution

iii. 
F = 20.438 

p-value = 0.000169

p-value is very small, much less than .05 and F is quite large; we do reject the null hypothesis and conclude that x2, passing yardage, is a significant contributor.

### (h)
```{r}
#investigate the contribution of the 3 predictors
reduced1 <- lm(y~x2+x8, data = ftbl)
anova(reduced1, lm1)
```
i. 

H0: βˆ1 = 0, βˆ5 = 0, βˆ7 = 0

H1: βˆ1 != 0, βˆ5 != 0, βˆ7 != 0

ii. F-distribution

iii. 

F = 1.6144

p-value = 0.2147

p-value is not less than .05 and F is quite small; we do not reject the null hypothesis and conclude that x1, x5, and x7 are not significant contributors.

### (i)
I prefer the reduced model with predictors x2 and x8 as it has a higher RSS and is more significant.

### (j)
R code is within this notebook pdf
