---
title: "STAT4355HW8"
author: "Ramesh Kanakala"
output: html_notebook
---

### Problem 1
### (a)
```{r}
#load the data
wind <- read.csv(file = 'windmill.csv')

#Box-Cox transformation
library(car)
boxcox.wind <- boxCox(wind$y ~ wind$x, 
                         lambda=seq(-2,2,1/10))
boxcox.wind # y=log-Likelihood, x=lambda
boxcox.wind$x[which.max(boxcox.wind$y)]
```

### (b)
```{r}
#build linear model
lm1 <- lm(y~x, wind)

#scatterplot with regression line
plot(wind$x, wind$y) + abline(lm1) 

#normal probability plot
qqPlot(lm1)

#studentized residuals versus the fitted values plot
par(mfrow=c(1,1))
residualPlot(lm1, type="rstudent", quadratic=F, col = "dodgerblue",
             pch=16, cex=1.5, cex.axis=1.5, cex.lab=1.5)

#influential analysis
myInf <- influence.measures(lm1)
myInf
summary(myInf)

dfbetasPlots(lm1,intercept=T)
influenceIndexPlot(lm1)
```

### Problem 2
### (a)
```{r}
#load the data
patient <- read.table('patient.txt', header=T)
x1 <- patient[,2] # age
x2 <- patient[,3] # severity
x3 <- patient[,4] # surgical-medical
x4 <- patient[,5] # anxiety
y <- patient[,6]  # satisfaction

x11=x12=x13=x14=array(0,length(y)) # n by 1 vector of 0's
x11[x1<30] <- 1          # if age group 1
x12[30<=x1 & x1<40] <- 1 # if age group 2
x13[40<=x1 & x1<50] <- 1 # if age group 3
x14[50<=x1 & x1<60] <- 1 # if age group 4

fit <- lm(y~x11+x12+x13+x14+x2+x3+x4)

#test for a general linear hypothesis
library(multcomp)
D <- matrix(0,3,8)
D[1,2]=D[2,3]=D[3,4]=1
D

d <- c(0,0,0)
mytest <- glht(fit, linfct=D, rhs=d)  # from page 15
summary(mytest,test=Ftest())
```

i. 

H0: β11 = β12 = β13 = 0

H1: β11 or β12 or β13 /= 0

ii.

D =
01000000
00100000
00010000

d = 
000

iii.

As F = 11.78488 and p = 0.0002035494 < 0.05, we reject the null hypothesis and confirm either β11 or β12 or β13 does not equal zero.

### (b)
```{r}
x11=x12=array(0,length(y))
x11[x1<50] <- 1         # if age group 1
x12[50<=x1 & x1<60] <- 1 # if age group 2

fit <- lm(y~x11+x12+x2+x3+x4)
summary(fit)
```
x11: With all other variables remaining the same, if a patient is in the age group under 50, satisfaction is associated with a 48.8871 increase on average.

x12: With all other variables remaining the same, if a patient is in the age group between 50 and 60, satisfaction is associated with a 32.6479 increase on average.

x2: With all other variables remaining the same, for a one unit increase in severity, satisfaction is associated with a 0.4358 decrease on average.

x3: With all other variables remaining the same, if a patient is a medical patient, satisfaction is associated with a 1.7310 decrease on average.

x4: With all other variables remaining the same, for a one unit increase in anxiety, satisfaction is associated with a 4.9538 increase on average.

### R code is within this notebook

